---
title: "ISyE 3770 HW5"
author: "Devin Fromond"
date: "2024-10-06"
output:
  pdf_document: default
  html_document: default
---

## Problem 5.1.4

### Part A:

We observe this problem to resemble a hypergeometric distribution, with a PMF as follows:

$$
f_{XY}(x,y)=\frac{\binom{30}{x} \binom{10}{y} \binom{60}{4-x-y}}{\binom{100}{4}}
$$

Therefore, we have the following table:

\[
\begin{array}{|c|c|c|c|c|c|c|}
\hline
X | Y & 0 & 1 & 2 & 3 & 4 & \text{Total}\ f_X(x) \\
\hline
0 & 0.1244 & 0.0873 & 0.02031 & 0.0018 & 0.0001 & 0.234 \\
1 & 0.2618 & 0.13542 & 0.02066 & 0.00092 & 0 & 0.419 \\
2 & 0.1964 & 0.0666 & 0.00499 & 0 & 0 & 0.268 \\
3 & 0.0621 & 0.01035 & 0 & 0 & 0 & 0.073 \\
4 & 0.0069 & 0 & 0 & 0 & 0 & 0.007 \\
\hline
\text{Total}\ f_Y(y) & 0.6516 & 0.2996 & 0.0460 & 0.0028 & 0.0001 & 1 \\
\hline
\end{array}
\]

### Part B:

\[
\begin{array}{|c|c|c|c|c|c|}
\hline
X & 0 & 1 & 2 & 3 & 4 \\
\hline
f_X(x) & 0.234 & 0.419 & 0.268 & 0.073 & 0.007 \\
\hline
\end{array}
\]

### Part C: 

$$
E[X]=]\sum^4_{x=0}xf_X(x)=0(0.234)+1(0.419)+2(0.268)+3(0.073)+4(0.007)
$$

```{r webSiteA}
0 * 0.234 + 1 * 0.419 + 2 * 0.268 + 3 * 0.073 + 4 * 0.007
```

### Part D:

We know the following:

$$
P(A|B)=\frac{P(A,B)}{P(B)}
$$
Therefore, we have

$$
f_{Y|X=3}(y)=\frac{f_{XY}(3,y)}{f_X(3)}=\frac{f_{XY}(3,y)}{0.0725}
$$

And so, we arrive at our answer of

\[
f_{Y|X=3}=\begin{array}{|c|c|c|c|c|c|}
\hline
X & 0 & 1 & 2 & 3 & 4 \\
\hline
f_X(x) & 0.857 & 0.143 & 0 & 0 & 0 \\
\hline
\end{array}
\]

### Part E: 

$$
E[Y|X=3]=\sum^4_{y=0}yf_Y(y|X=3)=0(0.857)+1(0.143)+0+0+0
$$

```{r webSiteE}
0 * 0.857 + 1 * 0.143 + 0 + 0 + 0
```

### Part F:

$$
V(Y|X=3)=0^2(0.857)+1^2(0.143)-(0.143)^2=0.123
$$

### Part G:

For independence, we know $f_{XY}(x,y)=f_X(x)f_Y(y)$. Therefore, we find the following:

$$
f_Y(2)\neq0
$$
$$
f_{Y|X=3}(2)=0,\ as\ previously\ found
$$

Therefore, since the previous two equations are not equal, they are not independent. 

## Problem 5.1.9

$$
\int^3_0\int^3_0cxydxdy=\int^3_0(cy\frac{x^2}{2})|^3_0dy=\int^3_2\frac{9}{2}cydy=(\frac{9cy^2}{4})|^3_0=\frac{81c}{4}=1
$$

Therefore, we have $c=\frac{4}{81}$.

### Part A:

$$
P(X<2, Y<2)=\int^3_0\int^2_0\frac{4}{81}xydxdy=\int^3_0(\frac{4yx^2}{162})|^2_0dy=(\frac{81y^2}{162})|^3_0=\frac{4}{9}=0.4444
$$

### Part B: 

$$
P(X<2.5)=\int^3_0\int^{2.5}_0\frac{4}{81}xydxdy=\int^3_0(\frac{4yx^2}{162})|^{2.5}_0dy=(\frac{12.5y^2}{162})|^3_0=\frac{6.25}{9}=0.694444
$$

### Part C: 

$$
P(1<Y<2.5)=\int^{2.5}_1\int^3_0=\int^{2.5}_1(\frac{4yx^2}{162})|^3_0dy=(\frac{2y^2}{18})|^{2.5}_{1}=\frac{7}{12}=0.5833
$$

### Part D: 

$$
P(X>1.8,1<Y<2.5)=\int^{2.5}_1\int^3_{1.8}\frac{4}{81}xydxdy=\int^{2.5}_1\frac{32y}{225}dy=\frac{28}{75}=0.3733
$$


### Part E:

$$
E[X]=\int^3_0xf_X(x)dx=\int^3_0x\frac{2x}{9}=(\frac{2x^3}{27})|^3_0=2
$$

### Part F:

$$
P(X<0,Y<4)=0,\ since\ X\ is\ restricted\ to\ 0<x<3
$$

### Part G:

$$
f_X(x)=\int^3_0\frac{4}{81}xydy=(\frac{2xy^2}{81})|^3_0=\frac{2x}{9},\ where\ 0<x<3
$$

## Problem 5.4.2

$$
\sum^3_{x=1}\sum^3_{y=1}c(x+y)=\sum^3_{x=1}c((x+1)+(x+2)+(x+3))=\sum^3_{x=1}c(3x+6)=3(1+2+3)+3(6)=c(18+18)=c36=1
$$
Therefore, we find $c=\frac{1}{36}$.

Next, we calculate Covariance.

$$
E[XY]=\sum^3_{x=1}\sum^3_{x=1}xy(\frac{1}{36})(x+y)
$$

\textbf{For \(x = 1\):}
\begin{align*}
y = 1: & \quad 1 \cdot 1 \cdot \frac{1}{36}(1 + 1) = \frac{2}{36} \\
y = 2: & \quad 1 \cdot 2 \cdot \frac{1}{36}(1 + 2) = \frac{6}{36} \\
y = 3: & \quad 1 \cdot 3 \cdot \frac{1}{36}(1 + 3) = \frac{12}{36} 
\end{align*}

\textbf{For \(x = 2\):}
\begin{align*}
y = 1: & \quad 2 \cdot 1 \cdot \frac{1}{36}(2) = \frac{4}{36} \\
y = 2: & \quad 2 \cdot 2 \cdot \frac{1}{36}(2 + 2) = \frac{16}{36} \\
y = 3: & \quad 2 \cdot 3 \cdot \frac{1}{36}(2 + 3) = \frac{30}{36} 
\end{align*}

\textbf{For \(x = 3\):}
\begin{align*}
y = 1: & \quad 3 \cdot 1 \cdot \frac{1}{36}(3 + 1) = \frac{12}{36} \\
y = 2: & \quad 3 \cdot 2 \cdot \frac{1}{36}(3 + 2) = \frac{30}{36} \\
y = 3: & \quad 3 \cdot 3 \cdot \frac{1}{36}(3 + 3) = \frac{54}{36} 
\end{align*}

$$
E[XY] = \frac{2}{36} + \frac{6}{36} + \frac{12}{36} + \frac{4}{36} + \frac{16}{36} + \frac{30}{36} + \frac{12}{36} + \frac{30}{36} + \frac{54}{36}
$$

$$
E[XY]=4.66666
$$
$$
E[X]=\sum^3_{x=1}xf_X(x)=\frac{1}{4}+\frac{2}{3}+\frac{5}{4}=\frac{13}{6}
$$
By symmetry, $E[Y]=E[X]=\frac{13}{6}$
$$
Cov(X,Y)=4.66666-\frac{169}{36}=-0.0277
$$

Now, we calculate Correlation:
$$
\rho(X,Y)=\frac{-0.0277}{\sqrt{\frac{23}{36}\cdot\frac{23}{36}}}=-0.04347
$$

## Problem 5.4.6
We first calculate covariance.
$$
f_X(x)=\int^\infty_0e^{-x-y}dy=e^{-y}\int^\infty_0e^{-x}dx=e^{-x}
$$
$$
f_Y(y)=e^{-y}
$$
$$
E[X]=\int^\infty_0xe^{-x}dx=(-xe^{-x})|^\infty_0+\int^\infty_0e^{-x}dx=0+1=1=E[Y]
$$
$$
E[XY]=(\int^\infty_0xe^{-x}dx)(\int^\infty_0ye^{-y}dy)=(1)(1)=1
$$
$$
Cov(X,Y)=E[XY]-E[X]E[Y]=1-(1)(1)=0
$$
Now, we will calculate correlation. 
$$
Var(X)=E[X^2]-(E[X])^2=\int^\infty_0x^2e^{-x}dx-1^2=2-1^2=1=Var(Y)
$$
$$
\rho=\frac{Cov(X,Y)}{\sqrt{Var(X)Var(Y)}}=\frac{0}{1\cdot 1}=0
$$