---
title: "ISyE 3770 HW10"
author: "Devin Fromond"
date: "2024-12-04"
output:
  pdf_document: default
  html_document: default
---
## Problem 10.2.6

### Part A

We will utilize Welch's t-test. 

We begin by calculating the standard error:
$$
SE=\sqrt{\frac{12^2}{10}+\frac{22^2}{16}}=\sqrt{44.65}=6.682
$$

Then, we determine the test statistic: 
$$
t=\frac{321-290}{6.682}=4.64
$$

Now, we use R to speed up calculations:
```{r partA}
n1 <- 10
s1 <- 12
n2 <- 16
s2 <- 22

numerator <- ( (s1^2 / n1) + (s2^2 / n2) )^2

denominator <- ( ( (s1^2 / n1)^2 ) / (n1 - 1) ) + ( ( (s2^2 / n2)^2 ) / (n2 - 1) )

df <- numerator / denominator

df

t_value <- 4.64
p_val <- 1- pt(t_value, df)
p_val
```

Therefore, since we have a p-value of $0.00005 > 0.05 = \alpha$, we reject the null hypothesis. Therefore, we find there is strong evidence to support the claim that Supplier 2 provides more gears with higher mean impact strength. 

### Part B

We again calculate the test statistic:
$$
t=\frac{(321-290)-25}{6.682}=0.898
$$

```{r partB}
t_value <- 0.898
df <- 23.72113
p_value <- 1 - pt(t_value, df)
p_value
```

Since the p-value of $0.1891>0.05=\alpha$, we fail to reject the null hypothesis. Therefore, we find the data does not support the mean impact strength of gears from Supplier 2 is at least 25 ft-lbs higher than that of Supplier 1. 

### Part C

We construct a 95% CI:

```{r partC}
df <- 23.72113
alpha <- 0.05
t_star <- qt(1 - alpha/2, df)
t_star
```

We then calculate the margin of error:
$$
ME=t*(SE)=2.065(6.682)=13.79833
$$

Therefore, we observe a CI of:
$$
(321-290) \pm 12.79 = 31 \pm 13.79
$$
This gives us a lower limit of $17.21$ and an upper limit of $44.79$. Therefore, we find the 95% CI to be $(17.21,44.79)$. 

Since the entire interval is $>0$, it supports the conclusion from Part A. Additionally, since the lower limit is $<25$, it supports the conclusion in Part B. 

The CI suggests the true difference is between 17.21 and 44.79 ft-lbs. 