---
title: "ISyE 3770 HW1"
author: "Devin Fromond"
date: "2024-09-03"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

## Problem 2.1.13

Using the Region Numbers as outlined in class, we have:

- **Part (a) \(A'\)**: First, we identify \(A\) as the green circle. Since part (a) asks for the elements not in \(A\), the answer is: \(\text{3 + 5 + 6 + 8}\).

- **Part (b) \(A \cap B\)**: First, we find the elements that exist in both \(A\) and \(B\). Therefore, we find the answer to be: \(\text{1 + 2}\).

- **Part (c) \((A \cap B) \cup C\)**: We begin by identifying the intersection of \(A\) and \(B\), which we see to be \(1+2\) as previously found. Now, we will find the elements that are in \(C\) and add them to the list. Finally, we have the following as our answer: \(\text{1+2+3+4+5}\).

- **Part (d) \((B \cup C)'\)**: We will start by identifying the union of \(B\) and \(C\) to be \(1+2+3+4+5+6\). Now, we will find the answer not to be \(\text{7+8}\). 

- **Part (e) \((A \cap B)' \cup C\)**: First, we will identify \(A\) intersection \(B\) or, the elements both \(A\) and \(B\) have in common. We find this to be \(1+2\). Then, we will find answer not to be \(3+4+5+6+7+8\). Then, we will find the answer union \(C\) to be \(\text{1+3+4+5+6+7+8}\). 

## Problem 2.1.21
Using the definitions of \(A\) and \(B\), we have the following:

- **Part (a) \(A'\)**: We begin by identifying \(A\) to be elements less than \(72.5\). Therefore, we find \(A'\) to be elements greater than or equal to \(72.5\)}. Symbolically, we have: $\text{\(A'=\{x\ |\ x \geq 72.5\}\)}$.

- **Part (b) \(B'\)**: We find \(B\) to be elements greater than \(52.5\). Therefore, we find \(B'\) to be elements less than or equal to \(52.5\). Symbolically, we have $\text{\(B'=\{x\ |\ x \leq 52.5\}\)}$.

- **Part (c) \(A \cap B\)**: The intersection of \(A\) and \(B\) include elements in both \(A\) and \(B\). Therefore, we find such elements to be those between \(52.5\) and \(72.5\), exclusive. Symbolically, we have $\text{\(A \cap B=\{x\ |\ 52.5 < x <72.5\}\)}$.

- **Part (d) \(A \cup B\)**: The union of \(A\) and \(B\) include elements that are in \(A\) or \(B\) (or both). Therefore, we find such elements to be those less than \(72.5\) and those greater \(52.5\). Symbolically, we have $\text{\(A \cup B=\{x\ |\ x > 0\}\)}$.

## Problem 2.2.13

We must find samples containing at least four defective parts, so we must find two answers: samples containing \(4\) defective parts and samples containing \(5\) defective parts.

We recognize order does not matter, so we proceed using a combination. 

For \(4\) defective parts, we know we must pick four of the five defective parts then we must pick six of the forty-five other parts. Therefore, we have

\[
C^5_4 C^{45}_6
\]

For \(5\) defective parts, we follow similar logic:

\[
C^5_5 C^{45}_5
\]

Therefore, we find our total number of samples to contain at least four defective parts to be \(C^5_4 C^{45}_6 + C^5_5 C^{45}_5\) = 40,725,300 + 1,221,759 = $\text{41,947,059}$.

## Problem 2.3.1

- **Part (a) \(P(A)\)**: We know the probability of an event is the sum of the probabilities of the outcomes in the event. As such, we have:

\[
P(A) = P({a, b, c}) = P({a}) + P({b}) + P({c}) = 0.1 + 0.1 + 0.2 = \text{0.4}
\]

- **Part (b) \(P(B)\)**: Following the previous question's logic, we have:

\[
P(B) = P({c, d, e}) = P({c}) + P({d}) + P({e}) = 0.2 + 0.4 + 0.2 = \text{0.8}
\]

- **Part (c) \(P(A')\)**: We remember \(P(A)=1-P(A)\) and so, we have our answer:

\[
P(A') = 1 - P(A) = 1 - 0.4 = \text{0.6}
\]

- **Part (d) \(P(A \cup B)\)**: Since the union of \(A\) and \(B\) cover all events in the sample space, we arrive at our answer of \(P(A \cup B) = \text{1}\).

- **Part (e) \(P(A \cap B)\)**: There exists only one element in the intersection of \(A\) and \(B\) and so, we arrive at our answer of \(P(A \cap B) = P(\{c\})= \text{0.2}\).

## Problem 2.3.17

- **Part (a) \(P(A)\)**: The probability of \(A\) is the probability of a password, containing 8 characters, being composed of only letters. For the first character, we know there are 62 total choices of characters, but 52 that are a letter. We will use this logic 8 times, so we have:

\[
P(A) = \left(\frac{52}{62}\right)^8 \approx 0.24485
\]



- **Part (b) \(P(B)\)**: The probability of \(B\) is the probability of a password, containing 8 characters, being composed of only integers. Using similar logic as previously, we have:

\[
P(B) = (\frac{10}{62})^8 = \text{4.58*10\(^{-7}\)}
\]

- **Part (c) \(P(Password\ containing\ at\ least\ 1\ integer)\)**: The probability of a password containing exactly 1 integer follows \(P(A')=1-P(A)\), so we have \(P(A')=1-0.24485=\text{0.75515}\).

- **Part (d) \(P(Password\ containing\ exactly\ 2\ integers)\)**: We wish for two of the 8 positions in the password to be an integer. Therefore, we have \(C^8_2=28\). For the two selected positions, there exist 10 possible integers, or 100 possible characters. For the remaining 6 positions, we have 52 options, resulting in \(52^6\) possibilities. Therefore, the number of passwords containing exactly two integers is:
\[
28*100*52^6
\]
Therefore, we arrive at our answer of \(\frac{28*100*52^6}{62^8}=\text{0.2535}\).

## Problem 2.4.7

- **Part (a) \(P(A)\)**: Since \(P(A)\) represents the probability of selecting a password beginning with a vowel, we know we \(C^5_1\) represents choosing a vowel to start the password and, to fill the remaining 5 characters, there are \(36^5\) options. Therefore, we have the following:

\[
P(A) = \frac{C^5_1*36^5}{36^6}=\text{\(\frac{5}{36}\)}
\]


- **Part (b) \(P(B)\)**: Abstracting this problem, it is asking the probability of one of five characters being in a particular position of the password... which is the same as **part (a)**. As such, we have:

\[
P(B) = \frac{C^5_1*36^5}{36^6}=\text{\(\frac{5}{36}\)}
\]

- **Part (c) \(P(A \cap B)\)**: Using the intersection rule of probability, we have \(P(A \cap B)=P(A)*P(B)\). We recognize events \(A\) and \(B\) are independent of each other - the last character being an even number does not influence the probability of the first being a vowel. As such, we have: 

\[
P(A \cap B) =P(A)*P(B) = \frac{5}{36}*\frac{5}{36}=\text{\(\frac{25}{1296}\)}
\]


- **Part (d) \(P(A \cup B)\)**: Using the union rule of probability, we have \(P(A \cup B)=P(A)+P(B)-P(A \cap B)\). As such, we have the following:

\[
P(A \cup B) =\frac{5}{36} + \frac{5}{36} - \frac{25}{1296} = \text{0.2584}
\]

## Question 2: Poker
To find the number of combinations for choosing three of the same cards, we must fist choose a rank, represented by \(C^{13}_1\). Then, we must choose three of the four cards of the rank: \(C^4_3\). We must then select one of the remaining 12 ranks and choose one card: \(C^{12}_1C^4_1\). Then, we must do this once more with the remaining 11 ranks: \(C^{11}_1C^4_1\). Finally, we have:

\[
C^{13}_1*C^4_3*C^{12}_1*C^4_1*C^{11}_1*C^4_1=
\]

```{r Poker Calculation, echo = FALSE }
choose(13,1)*choose(4,3)*choose(12,1)*choose(4,1)*choose(11,1)*choose(4,1)
```

## Question 3: Four Candidates

This question follows very similar to a repeated combination example. In this case, we have \(n=4\) and \(r=100\), or a selection of 100 items with repetition from a set of 4. As such, we have:
\[
C^{n+r-1}_r=C^{4+100-1}_{100}=
\]

```{r Four Candidates, echo = FALSE}
choose(103, 100)
```

