---
title: "ISyE 3770 HW6"
author: "Devin Fromond"
date: "2024-10-24"
output:
  pdf_document: default
  html_document: default
---

## Problem 6.1.6

We observe the sample mean is the average across all data points:

$$
\bar{x}=\frac{\sum^n_{i=1}x_i}{n}
$$

Therefore, we have the following:

```{r Mean}
0.19 + 0.78 + 0.96 + 1.31 + 2.78 + 3.16 + 4.15 + 4.67 + 4.85 + 6.50 + 7.35 + 8.01 + 8.27 + 12.06 + 31.75 + 32.52 + 33.91 + 36.71 + 72.89

```

$$
\bar{x}=\frac{272.82}{19}=14.359
$$

We observe the sample standard distribution is defined as the following:

$$
s=\sqrt{\frac{\sum^n_{i=1}(x_i-\bar{x})^2}{n-1}}
$$

We will use R to speed up the process...
```{r STD DEV}
breakdown_times <- c(0.19, 0.78, 0.96, 1.31, 2.78, 3.16, 4.15, 4.67, 4.85, 6.50, 
                    7.35, 8.01, 8.27, 12.06, 31.75, 32.52, 33.91, 36.71, 72.89)

variance <- sum((breakdown_times - 14.359)^2) / 18
sqrt(variance)
```


## Problem 6.2.4


Let's start with the stem-and-leaf plot:
```{r OctaneLeaf}
octane_ratings <- c(
  88.5, 98.8, 89.6, 92.2, 92.7, 88.4, 87.5, 90.9,
  94.7, 88.3, 90.4, 83.4, 87.9, 92.6, 87.8, 89.9,
  84.3, 90.4, 91.6, 91.0, 93.0, 93.7, 88.3, 91.8,
  90.1, 91.2, 90.7, 88.2, 94.4, 96.5, 89.2, 89.7,
  89.0, 90.6, 88.6, 88.5, 90.4, 84.3, 92.3, 92.2,
  89.8, 92.2, 88.3, 93.3, 91.2, 93.2, 88.9,
  91.6, 87.7, 94.2, 87.4, 86.7, 88.6, 89.8,
  90.3, 91.1, 85.3, 91.1, 94.2, 88.7, 92.7,
  90.0, 86.7, 90.1, 90.5, 90.8, 92.7, 93.3,
  91.5, 93.4, 89.3, 100.3, 90.1, 89.3, 86.7,
  89.9, 96.1, 91.1, 87.6, 91.8, 91.0, 91.0
)

stem(sort(octane_ratings), scale = 2)
```

Now, we can calculate the median and quartiles:

```{r OctaneMedandQuart}
median_octane <- median(octane_ratings)
quartiles <- quantile(octane_ratings, type=6)

median_octane
quartiles
```

## Problem 6.3.2

```{r frequency-distribution}
library(dplyr)
library(knitr)

octane_ratings <- c(
  88.5, 98.8, 89.6, 92.2, 92.7, 88.4, 87.5, 90.9,
  94.7, 88.3, 90.4, 83.4, 87.9, 92.6, 87.8, 89.9,
  84.3, 90.4, 91.6, 91.0, 93.0, 93.7, 88.3, 91.8,
  90.1, 91.2, 90.7, 88.2, 94.4, 96.5, 89.2, 89.7,
  89.0, 90.6, 88.6, 88.5, 90.4, 84.3, 92.3, 92.2,
  89.8, 92.2, 88.3, 93.3, 91.2, 93.2, 88.9,
  91.6, 87.7, 94.2, 87.4, 86.7, 88.6, 89.8,
  90.3, 91.1, 85.3, 91.1, 94.2, 88.7, 92.7,
  90.0, 86.7, 90.1, 90.5, 90.8, 92.7, 93.3,
  91.5, 93.4, 89.3, 100.3, 90.1, 89.3, 86.7,
  89.9, 96.1, 91.1, 87.6, 91.8, 91.0, 91.0
)

num_bins <- 8
breaks <- seq(floor(min(octane_ratings)), ceiling(max(octane_ratings)), length.out = num_bins + 1)
hist_result <- hist(octane_ratings, breaks = breaks, plot = FALSE)
freq <- hist_result$counts

bins <- paste0(head(breaks, -1), " \\( \\leq x < \\) ", tail(breaks, -1))

bins[length(bins)] <- paste0(breaks[num_bins], " \\( \\leq x \\leq \\) ", breaks[num_bins + 1])

relative_freq <- freq / sum(freq)
cumulative_relative_freq <- cumsum(relative_freq)

freq_table <- data.frame(
  Class = bins,
  Frequency = freq,
  `Relative Frequency` = round(relative_freq, 4),
  `Cumulative Relative Frequency` = round(cumulative_relative_freq, 4)
)

kable(freq_table, caption = "Frequency Distribution of Motor Fuel Octane Ratings")

hist(octane_ratings, breaks = breaks, col = "skyblue", border = "black",
     main = "Histogram of Motor Fuel Octane Ratings",
     xlab = "Octane Rating",
     ylab = "Frequency")
```

## Problem 6.4.9

```{r bigData}
library(ggplot2)

data <- read.table("C:/Users/devin/OneDrive/Documents/Georgia_Tech/6-81.txt", header = TRUE)

ggplot(data, aes(x = Group, y = Expression)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  labs(title = "Gene Expression for Different Treatment Groups",
       x = "Group",
       y = "Gene Expression") +
  theme_minimal()
```

The HighDose displays lower gene expression levels than the control groups. 
The control groups demonstrate greater variability in gene expression; Control3 has multiple outliers which skew the data.

## Problem 6.7.2

```{r MORE OCTANE}
library(ggplot2)

ggplot(data.frame(octane_ratings), aes(sample = octane_ratings)) +
  stat_qq() +
  stat_qq_line(color = "red") +
  labs(title = "Normal Prob Plot of Octane Ratings",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme_minimal()
```

Points which lie approximately on the red line indicate a normal distribution. 

Points which deviate from the red line suggest deviations from a normal distribution. 

Since most of the points lie approximately on the red line, we can conclude the octane rating is normally distributed. 