---
title: "ISyE 3770 HW8"
author: "Devin Fromond"
date: "2024-11-11"
output:
  pdf_document: default
  html_document: default
---

## Problem 8.1.1

### Part A

Given the margin of error is $2.14\frac{\sigma}{\sqrt{n}}$, we easily see $z=2.14$. 

Using a Z-Table, we find the following:
$$
P(Z\leq 2.14)\approx 0.9838
$$
$$
P(Z \leq -2.14)\approx 1-0.9838=0.0162
$$
$$
Confidence\ Level=0.9838-0.0162=0.9676\ or\ 96.76%
$$

### Part B

We observe a Z-Score of $2.49$.

Using a Z-Table, we find the following:
$$
P(Z\leq 2.49)\approx 0.9934
$$
$$
P(Z \leq -2.49)\approx 1-0.9934=0.0066
$$
$$
Confidence\ Level=0.9934-0.0066=0.9868\ or\ 98.68%
$$

### Part C

We observe a Z-Score of $1.85$.

Using a Z-Table, we find the following:
$$
P(Z\leq 1.85)\approx 0.9678
$$
$$
P(Z \leq -1.85)\approx 1-0.9678=0.0322
$$
$$
Confidence\ Level=0.9678-0.0322=0.9356\ or\ 93.56%
$$

### Part D

We recognize this is a one-sided interval. We observe a Z-Score of $2.00$.

Using a Z-Table, we find the following:
$$
P(Z\leq 2.00)\approx 0.9772
$$
$$
Confidence\ Level=P(\mu \leq \bar{x}+2.00\frac{\sigma}{\sqrt{n}})=P(Z\leq2.00)=0.9772\ or\ 97.72%
$$

### Part E

We recognize this is a one-sided interval. We observe a Z-Score of $1.96$.

Using a Z-Table, we find the following:
$$
P(Z\leq 1.96)\approx 0.9750
$$
$$
Confidence\ Level=P(\mu \leq \bar{x}-1.96\frac{\sigma}{\sqrt{n}})=P(Z\leq1.96)=0.9750\ or\ 97.50%
$$

## Problem 8.1.5

### Part A

A higher confidence interval would take longer. Therefore, we find that a 99% CI would be longer than a 95% CI since the interval must be wider to account for greater variability in the sample data.

### Part B

The statement is not correct. For a CI $X%$, we find $X%$ of samples of those intervals would contain the true mean $\mu$. Furthermore, we find $\mu$ is not a random variable, so it is not within a specific interval. The provided statement treats $\mu$ as a random variable.

### Part C

We find the statement to be correct. The statement correctly interprets confidence intervals, as outlined in Part B. Out of 1000 confidence intervals, we would expect $\approx 950$ to include $\mu$, which we find to be consistent with the CI. 

## Problem 8.1.7

### Part A

We find $Z_{0.005}\approx 2.5758$. 

Therefore, we calculate standard error to be:
$$
SE=\frac{\sigma}{\sqrt{n}}=\frac{0.001}{\sqrt{15}}=0.0002582mm
$$

We compute the margin of error to be:
$$
ME=Z_{0.005}(SE)=2.5758(0.0002582)=0.000664
$$

Therefore, we find the CI to be:
$$
CI=(74.036-0.000664, 74.036 + 0.000)=(74.0353mm,\ 74.0367mm)
$$

### Part B
We find $Z_{0.99}\approx 2.3263$. 
Therefore, we calculate standard error to be:
$$
SE=\frac{\sigma}{\sqrt{n}}=\frac{0.001}{\sqrt{15}}=0.0002582mm
$$

We compute the margin of error to be:
$$
ME=Z_{0.99}(SE)=2.3263(0.0002582)=0.000600
$$

Therefore, we find the Lower Confidence Bound to be:
$$
LCB=74.036-0.000600\approx 74.0354mm
$$

We find the lower bound of the one-sided 99% confidence bound is slightly greater than that of the two-sided 99% confidence interval. We find this to be logical since the two-sided interval divides the significance level to each tail, thereby making it more conservative. 

## Problem 8.2.10

### Part A

```{r 8.2.10A}
speedup <- c(3.775302, 3.350679, 4.217981, 4.030324, 4.639692,
            4.139665, 4.395575, 4.824257, 4.268119, 4.584193,
            4.930027, 4.315973, 4.600101)

qqnorm(speedup, main="Q-Q Plot of CNN Speedup")
qqline(speedup, col="red", lwd=2)
```

Since the data points in the Q-Q plot lie approximately on the red line, we can conclude the speedup of CNN is normally distributed. 

### Part B

We begin by calculating the sample mean:
$$
\bar{x}=\frac{\sum x_i}{n}=\frac{56.071888}{13}=4.3132
$$

We then calculate the sample standard deviation:
$$
SST=\sum (x_i-\bar{x})^2=2.2478
$$

Now, we find the variance to be:
$$
S^2=\frac{SST}{n-1}=\frac{2.2478}{12}=0.1873
$$

The standard deviation is trivial:
$$
S=\sqrt{S^2}=0.4328
$$

Therefore, we can calculate the standard error:
$$
SE=\frac{S}{\sqrt{n}}=\frac{0.4328}{\sqrt{13}}=0.1201
$$

We find the following:
$$
t_{\alpha /2, n-1}=t_{0.025,12}=2.179
$$

Therefore, we find the margin of error to be $2.179(0.1201)=0.2616$. 

Then, the lower limit is calculated by $4.3132-0.2616=4.0515$ and the upper bound is $4.3132+0.2616=4.5748$.

### Part C

Using the values from part B, we simply find the following:
$$
t_{\alpha , n-1}=t_{0.05,12}=1.782
$$

Therefore, we find the margin of error to be $1.782(0.1201)=0.2140$.

Finally, the lower bound is found by $4.3132-0.2140=4.0992$. 

## Problem 8.3.3

We begin by calculating the mean and standard deviation.
```{r mean_sd}
temperatures <- c(20.1, 20.7, 20.9, 21.3, 21.6, 21.7, 21.7, 23.3)

mean_temp <- mean(temperatures)

sd_temp <- sd(temperatures)

print(mean_temp)
print(sd_temp)
```

For each temperature, we calculate the z-score. 
```{r zscore}
z_scores <- (temperatures - mean_temp) / sd_temp

temperature_data <- data.frame(
  Site = c("Harlquahala", "Laveen", "Maricopa", "Litchfield",
           "Phoenix Encanto", "Phoenix Greenway", "Waddell", "Sky Harbor Airport"),
  Temperature_C = temperatures,
  Z_Score = round(z_scores, 3)
)

print(temperature_data)
```

Now, we plot the data!!!
```{r qq}
qqnorm(temperatures, main = "Q-Q Plot of Mean Annual Temperatures", pch = 19, col = "blue")
qqline(temperatures, col = "red", lwd = 2)

grid()
```

Since most of the data points lie on or around the red line, we confirm the assumption of normality. 

Now, we construct a 95% confidence interval. 

Let's use R to make this process easier...
```{r CI}
variance_temp <- var(temperatures)

df <- length(temperatures) - 1  
alpha <- 0.05  
chi_sq_lower <- qchisq(alpha / 2, df, lower.tail = FALSE)
chi_sq_upper <- qchisq(1 - alpha / 2, df, lower.tail = FALSE)

variance_lower <- (df * variance_temp) / chi_sq_lower
variance_upper <- (df * variance_temp) / chi_sq_upper

sd_lower <- sqrt(variance_lower)
sd_upper <- sqrt(variance_upper)
cat("95% Confidence Interval for Standard Deviation:\n")
cat("Lower Limit:", round(sd_lower, 3), "°C\n")
cat("Upper Limit:", round(sd_upper, 3), "°C\n")
```

## Problem 8.4.1

### Part A

Ultimately, we wish to solve the following to determine the confidence intervals:
$$
\hat{p} \pm Z_{\alpha/2}(\sqrt{\frac{\hat{p}(1-\hat{p})}{n}})
$$
, where $\hat{p}=\frac{x}{n}$.
We will use R to speed up calculations...

```{r Bush}
x <- 412 # bush voters
n <- 768 # college grads
conf_level <- 0.95

p_hat <- x / n

z <- qnorm(1 - (1 - conf_level)/2)

SE <- sqrt(p_hat * (1 - p_hat) / n)

l_bound <- p_hat - z * SE
u_bound <- p_hat + z * SE

cat("We find the 95% Confidence Interval for p to be", 
    round(l_bound, 5), "to", round(u_bound, 5))
```

### Part B

We find $\hat{p}=\frac{412}{768}=0.5365$. 

Therefore, we have:
$$
SE=\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}=0.0173
$$

Using R, 
```{r BushAgain}
z_lower <- qnorm(conf_level)

lower_conf_bound <- p_hat - z_lower * SE

cat("95% Lower Confidence Bound for p:", round(lower_conf_bound, 5))
```

And so, we have arrived at our answer of 50.69% for the 95% LCB.


