---
title: "ISyE 3770 HW4"
author: "Devin Fromond"
date: "2024-09-29"
output:
  pdf_document: default
  html_document: default
---

## Problem 3.8.12

### Part A:
$$
X \sim Pois(1.8)
$$
$$
P(X>5)=1-P(X \leq 5)=1-\sum^5_{k=0}\frac{e^{-1.8}1.8 ^k}{k!}
$$

```{r PoisA}
1 - sum(exp(-1.8) * 1.8^(0:5) / factorial(0:5))
```

### Part B:
$$
X \sim Pois(1.8\times 7)=Pois(12.6)
$$

$$
P(X<5)=\sum^4_{k=0}\frac{e^{-12.6}12.6 ^k}{k!}
$$
```{r PoisB}
sum(exp(-12.6) * 12.6^(0:4) / factorial(0:4))
```

### Part C:

$$
X \sim Pois(1.8\times d)=Pois(1.8d)
$$

$$
P(X \geq 1)=1-P(X < 1)=1-\frac{e^{-1.8d}(1.8d) ^0}{0!}=0.99
$$

$$
d=-\frac{ln(0.01)}{1.8}
$$

```{r PoisC}
(-1) * log(0.01) / 1.8
```

### Part D:

$$
X \sim Pois(m\times 1)=Pois(m)
$$
$$
P(X>5)=1-P(X \leq 5)=1-\sum^5_{k=0}\frac{e^{-m}m ^k}{k!}=0.1
$$

```{r PoisD}
uniroot(function(m) 1 - sum(exp(-m) * m^(0:5) / factorial(0:5)) - 0.1, lower = 0, upper = 10)$root

```

## Problem 4.1.2

### Part A:

$$
P(X<2)=\int^2_{1}\frac{2}{x^3}dx=\frac{3}{4}
$$

### Part B:

$$
P(X>5)=\int^\infty_{5}\frac{2}{x^3}dx=\frac{1}{25}
$$

### Part C:

$$
P(4<X<8)=\int^8_{4}\frac{2}{x^3}dx=\frac{3}{64}
$$

### Part D:

$$
P(X<4\ or\ X>8)=P(X<4)+P(X>8)=\int^\infty_{8}\frac{2}{x^3}dx+\int^4_{1}\frac{2}{x^3}dx=\frac{1}{64}+\frac{15}{16}=\frac{61}{64}
$$

### Part E:

$$
P(X<x)=\int^x_{1}\frac{2}{x^3}dx=0.95
$$

$$
(-\frac{1}{x^2}) \Big|_1^x=-\frac{1}{x^2}+1=0.95
$$
$$
x=\sqrt{20}
$$

## Problem 4.4.2

### Part A:

\[ F(x)= \begin{cases} 
      0 & x<0.95 \\
      \frac{x-0.95}{1.05-0.95}=10(x-0.95) & 0.95\leq x\leq 1.05 \\
      1 & x > 1.05 
   \end{cases}
\]

### Part B: 

$$
P(X>1.02)=1-F(1.02)=1-10(1.02-0.95)=1-0.7=0.3
$$

### Part C: 

$$
P(X>x)=0.9\ or\ F(x)=0.1
$$
$$
10(x-0.95)=0.1
$$

$$
x=0.96
$$

### Part D: 

$$
\mu=\frac{a+b}{2}=\frac{0.95+1.05}{2}=1.00
$$
$$
\sigma^2=\frac{(b-a)^2}{12}=\frac{(1.05-0.95)^2}{12}=\frac{(0.1)^2}{12}=0.000833
$$

## Problem 4.5.6

### Part A:


$$
Z=\frac{X-\mu}{\sigma}=\frac{240-260}{50}=-0.4
$$
$$
P(X>240)=1-P(Z<-0.4)=1-0.3446=0.6554
$$

### Part B: 
For the first Quartile:
$$
Z \approx -0.6745
$$
$$
X=\mu + Z(\sigma)=260+(-0.6745)(50)=226.275\ min
$$

For the second Quartile: 
$$
Z \approx 0.6745
$$
$$
X=260+(0.6745)(50)=293.7\ min
$$

### Part C: 

$$
Z \approx -1.645
$$
$$
X=260+(-1.645)(50)=177.75\ min
$$

## Problem 4.5.12

### Part A: 

$$
Z=\frac{X-\mu}{\sigma}=\frac{350-310}{45}=0.8889
$$

$$
P(Z>0.8889)=1-P(Z<0.8889)=0.1864
$$

### Part B:

$$
Z \approx 2.326
$$
$$
X=310+2.326(45)=414.67\ mill\ gal
$$

### Part C:

$$
Z \approx -1.645
$$
$$
X=310+(-1.645)(45)=235.975\ mill\ gal
$$

### Part D: 

$$
Z=\frac{X-\mu}{45}
$$
$$
2.33=\frac{350+\mu}{45}
$$
$$
\mu=245.2
$$

## Problem 4.7.5

### Part A: 

$$
P(T\leq30)=1-P(T>30)=1-1+e^{-\frac{1}{15}(30)}=0.1353
$$

### Part B: 

$$
P(T\leq 10)=1-e^{-\frac{1}{15}(10)}=0.4866
$$

### Part C: 

$$
P(5<T<10)=P(T<10)-P(T<5)=0.4866-(1-e^{-\frac{1}{15}(5)})=0.4866-0.2835=0.2031
$$

### Part D:

$$
P(T\leq t)=1-e^{-\lambda t}=0.90
$$
$$
1-e^{-\frac{1}{15}t}=0.90
$$
$$
t=34.54\ min
$$

## Problem 4.7.11

### Part A: 

$$
\lambda t=\frac{1}{10}(30)=3
$$
$$
N \sim Pois(3)
$$
$$
P(N>3)=1-P(N\leq 3)=1-\sum^3_{k=0}P(N=k)
$$
```{r callsA}
1 - (exp(-3) + 3 * exp(-3) + (9 * exp(-3)) / 2 + (27 * exp(-3)) / 6)
```

### Part B: 

As previously calculated:

$$
P(N=0)=e^{-3}=0.0498
$$

### Part C: 

$$
e^{-\frac{1}{10}(x)(60)}=0.01
$$
$$
x=\frac{4.605(10)}{6}=7.68\ hours
$$

### Part D:

$$
P(no\ calls\ in\ 120m)=e^{-\frac{1}{10}(120)}=0.00000614
$$

## Question 2

### Part A: CDF

$$
F(x)=\int^x_1\frac{2}{t^3}dt=[-\frac{1}{t^2}] \Big|^x_1=-\frac{1}{x^2}+1,\ for\ x>1
$$

\[ F(x)= \begin{cases} 
      0 & x\leq 1 \\
      1-\frac{1}{x^2} & x>1
   \end{cases}
\]


### Part B: Mean

$$
\mu=\int^\infty_1x(\frac{2}{x^3})dx=-2[\frac{1}{x}]\Big|^\infty_1=2
$$

### Part C: Variance

$$
Var(X)=E[X^2]-(E[X])^2
$$

$$
E[X^2]=\int^\infty_1x^2(\frac{2}{x^3})dx=\int^\infty_1\frac{2}{x}dx=[2ln(x)]\Big|^\infty_1=\infty
$$
Therefore, since $E[X^2]=\infty$, the variance is infinite. 